---
title: "DS311 - R Lab Assignment"
author: "Janaya"
date: "`r Sys.Date(today)`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)
dim(mtcars)
# Answer:
print("There are total of 32 variables and 11 observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are 0 discrete variables and 11 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!

m = mean(mtcars$mpg)
v = var(mtcars$mpg)
s = sd(mtcars$mpg)

# 
print(m)
print(v)
print(s)
print(paste("The average of Mile Per Gallon from this data set is ", round(m,2) , " with variance ", round(v,2) , "and standard deviation", round(s,2) "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!

data1 = setNames(aggregate(mpg~cyl, data=mtcars, FUN=mean), c('Cylinder', 'MPG_Avg'))

data2 = setNames(aggregate(mpg~gear, data=mtcars, FUN=mean), c('Gear', 'MPG_Stg'))

print(data1)
print(data2)
```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!

x = xtabs(~gear+cyl, data = mtcars)

print(x)
print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!

library(ggplot2)
g_plantgrowth = ggplot(data=PlantGrowth) + geom_boxplot(aes(x=group, y= weight))
g_plantgrowth = g_plantgrowth + labs(titale="PlantGRowth", x="grouping", y="weight")
g_plantgrowth

```

Result:

=> Report a paragraph to summarize your findings from the plot!

Trt2 has the highest plant weight and the most growth; therefore, it is the best result.

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
g_mtcars = ggplot(data=mtcars) + geom_histogram(aes(x=mpg)), colour="black", fill="white", binwidth = 10
g_mtcars = g_mtcars + lab(title = "mpg dist"), x="mpg", y="count"
g_mtcars

print("Most of the cars in this data set are in the class of ______ mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
library(ggplot2)

g_usarrests = ggplot(USArrests), aes(x=murder, y=assult, label=rownames(USArrests)) + geom_point()
g_usarrests = g_usarrests +labs(title="murder v assult", x="murder", y="assult")


```

Result:

=> Report a paragraph to summarize your findings from the plot!



***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
 count_neighborhood = housingData %>%
   grouping(neighborhood) %>% count()
 count_boro = housingData %>% grouping(boro) %>% count()
 count_builtyear = housingData %>% grouping(year.built) %>% count()
 count_neighborhood_boro_builtyear =  housingData %>% grouping(neighborhood, boro, yearbuilt) %>% count()

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!

library(ggplot2)
g_boro_built = ggplot(count_boro, aes(boro, n)) + geom_bar(stat="id")
g_boro_built = g_boro_built + labs(title='houses built in boro', x="noro", y="houses")
g_boro_built

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!


